The Astropy Project: Building an open-science project and status of the v2.0 core package

### 2018-02-28 main.tex
add context to new figure panel

As of version 2.0, \astropypkg contains $212244$ lines of code\footnote{This
line count includes comments, as these are often as important for
maintainability as the code itself.  Without comments there are $142197$ lines
of code.} contributed by $232$ unique contributors over $19270$ \texttt{git}
commits.
\figurename~\ref{fig:ncommits}, left, shows the distribution of total number of
commits per contributor as of November 2017.
The relative flatness of this distribution (as demonstrated by its log-log slope of
$-0.5$) shows that the \astropypkg core package has been developed by a broad
contributor base.  A leading group of 6 developers have each added over 1000
commits to the repository, and $\sim 20$ more core contributors have contributed
at least 100 commits.
However, the distribution of contribution level (number of commits) continues
from 100 down to a single commit.
In this sense, the development of the core package has been a true community
effort and is not dominated by a single individual.
It is also important to note that the number of commits is only a rough metric
of contribution, as a single commit could be a critical fix in the package or a
fix for a typographical error.
\figurename~\ref{fig:ncommits}, right, shows the number of commits as a
function of time since the genesis of the \astropypkg core package.
The package is still healthy: new commits are and have been contributed at a
steady rate throughout its existence.

@@

\subsection{Reuse of the scientific Python ecosystem}

The \astropy Project is built on a philosophy of building from the existing
Python scientific ecosystem wherever possible. Many of the enabling technologies
like \texttt{numpy}, \texttt{scipy}, \texttt{matplotlib}, or \texttt{cython},
are necessary as the core underpinnings of \astropy packages.  Often this means
simply means using these packages as ``building blocks'' (E.g., the ubiquitous
use of \texttt{numpy} arrays throughout \astropypkg).  In other cases this means
wrappers around more general algorithms that make them more convenient for
astronomy use cases (E.g., the model-fitting in the \package{astropy.modeling}
subpackage discussed in \sectionname~\ref{sec:modeling}).  Sometimes these
simple wrappers evolve into more complex implementations that address
astronomically-relevant use cases the general tool does not support (E.g.,
\package{astropy.convolution}, see \sectionname~\ref{ssec:convolution}). As a
broad rule, the Project explicitly encourages re-use of code where possible.

However, the boundaries of when this re-use is called for often ambiguous.  Some
of the examples above only evolved after significant debate in the community
over whether these algorithms were sufficient.  Other times, even apparently
general functionality did not exist in the wider ecosystem that met the \astropy
community's needs, and hence the functionality had to be developed wholly from
the developer resources available in the community (E.g.,
\package{astropy.units}, or \package{astropy.table} when it began). While some
of this functionality is ``given back'' to the wider community (E.g., myriad bug
fixes for the enabling technologies listed above, and some of the testing and
documentation architecture), in general, careful consideration is taken of
whether the additional effort to make a piece of \astropy, and it is only
``extracted'' when it seems plausible that the ongoing effort is worth the
reward in the form of additional contributors or maintainers.

@@

\subsection{Convolution}
\label{ssec:convolution}

>>>>

As of version 2.0, \astropypkg contains $212244$ lines of code\footnote{This
line count includes comments, as these are often as important for
maintainability as the code itself.  Without comments there are $142197$ lines
of code.} contributed by $232$ unique contributors over $19270$ \texttt{git}
commits.
\figurename~\ref{fig:ncommits}, left, shows the distribution of total number of
commits per contributor as of early 2018.
The relative flatness of this distribution (as demonstrated by its log-log slope of
$-0.5$) shows that the \astropypkg core package has been developed by a broad
contributor base.  A leading group of 6 developers have each added over 1000
commits to the repository, and $\sim 20$ more core contributors have contributed
at least 100 commits.
However, the distribution of contribution level (number of commits) continues
from 100 down to a single commit.
In this sense, the development of the core package has been a true community
effort and is not dominated by a single individual.
It is also important to note that the number of commits is only a rough metric
of contribution, as a single commit could be a critical fix in the package or a
fix for a typographical error.
\figurename~\ref{fig:ncommits}, middle, shows the number of commits as a
function of time since the genesis of the \astropypkg core package.
The package is still healthy: new commits are and have been contributed at a
steady rate throughout its existence.
\figurename~\ref{fig:ncommits}, right, shows that, $\approx$20--25 people
contribute to the core package each month.
While we would like for this number to grow, this demonstrates that the core
package is still being developed and maintained by a substantial group of
contributors.

@@

@@

\subsection{Convolution}
